---
title: "GlycoProteins"
author: "Maarten van Schaik"
date: "16 October 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

In this Markdown, I am exploring some options to display the results of multiple regression models in tabular format easily. This involves 3 steps:

1) Create some fake Glycan Peaks data (since I have no real data to show)
2) Use the `R` packages `tidyverse` and `broom` to efficiently calculate many models. This is highly preferred over making the models 'by hand' and copy-pasting code, because this automated way scales to any number of models, so running models over 1000 GPs is the same amount of code as over just 2 GPs.
3) Using the `stargazer` package to print out model results in a nice looking manner. It knows text, html and LaTeX. Later on, I might want to explore the `xtable` package as well (not yet done)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Required packages
require(tidyverse)
require(broom)
require(stargazer)
require(lattice) # for the plot in step 1
# require(xtable) # Maybe use later
# require(gridExtra) # to plot muliple panels together using grid.arrange()
```

# Create fake data

First, create some fake GP data and predictor data. I am just sampling from a couple of different gamma distributions to mimic GP distributions. I then normalize them so they all have mean 0 and sd 1. This makes it slightly easier to compare the distributions and the beta regression coefficients later on.

```{r}
# Create some fake outcomes using different Gamma distributions to mimic the shape 
# of real GP values
# However,they are completely independent of each other and the predictors
beta <- c(5, 4, 36, 40, 90, 35, 36, 35, 6.25, 4.5)
alpha <- c(0.5, 2, 6, 8, 2, 5, 6, 5, 2.5, 3)
N <- 1666
P <- 10
GP <- as.tbl(data.frame(matrix(NA, nrow=N, ncol=P)))
colnames(GP) <- paste0("GP", 1:10)
set.seed(1234)
for(i in 1:P){
  GP[,i] <- rgamma(n=N,
                   rate=alpha[i],
                   shape=beta[i])
}

# Normalise the data to make comparable across GPs. I also do this to avoid
# the problem of unequal x-ranges in histograms when plotting together using
# the lattice trellis plot
GP <- as.tbl(as.data.frame(scale(GP, center=TRUE, scale=TRUE)))

# Make into long format to use as input for lattice histogram trellis plot
GP_long <- GP %>%
  gather(GP, value, GP1:GP10)

# lattice trellis plot
lattice::histogram(~value | GP, 
                   data=GP_long,
                   scales="free")

```

```{r}
# Create fake predictors
X <- data.frame(id = factor(1:N),
                CTR_DIAB = factor(rbinom(n=N, size=1, p=0.5)),
                Complications = factor(rbinom(n=N, size=1, p=0.45)),
                Metabolic = factor(rbinom(n=N, size=1, p=0.25)),
                Age = rnorm(n=N, mean=58.5, sd=12.4),
                Sex = factor(rbinom(n=N, size=1, p=0.5)))

# Combine with fake GP outcomes
GP_wide <- as.tbl(cbind(X, GP))
```

# Proposed analysis (slow, by hand)

Do not run this code. It would be a way to (attempt to) extract model results for several different GP models using for loops and model summaries.

```{r, eval=FALSE}
## Table 2. Glycans ~ B1*CTR_DIAB + B2*Complications + Age + Sex
for (i in Glycan_Peaks.colonne){
  Taula.2 <- lm(Glycans[,i] ~ CTR_DIAB + Complications + eta + Sesso, Data_Gly_Comp)
  Vector_Peaks <- c(Vector_Peaks,lm(Taula.2))
  print(summary(Taula.2))
  print(summary(Taula.2)$coefficients)
  
  # How to extract the summary for each glycan peak?
  SE <- summary(Taula.2)[["coefficients"]][, "Std. Error"]
  P.value <- summary(Taula.2)[["coefficients"]][, "Pr(>|t|)"]
  #Taula.2.PValue <- data.frame(P.value)
  Taula.2.PValue[,i] <- data.frame(P.value)
  #Taula.2.SE <- data.frame(SE)
  Taula.2.SE[,i] <- data.frame(SE)
  # It doesn't work
}

```

# More efficiently, using packages

Now use the packages to run many models and extract results. 

First, combine the GP values and predictors in one table (already done above)
Then, we want to make a linear model for each GP, seperately. We can do this efficiently using map functions. For this, the data must be in long format first. 

```{r}
# Make long format
GP_long <- GP_wide %>%
  gather(GP, value, GP1:GP10) %>%
  arrange(id) 

GP_long

# Make nested
GP_nested <- GP_long %>%
  group_by(GP) %>%
  nest()

GP_nested # A tibble containing tibbles for each GP outcome variable
GP_nested$data[[1]] # The first GP-tibble. the 'values' correspond to GP1 values
```


```{r}
# Calculating the 10 linear models (one for each GP)

# For each of the 10 GPs, calculate a simple linear regression
GP_nested <- GP_nested %>%                               
    mutate(model = map(.x=data,                          
                       .f=~lm(value ~ CTR_DIAB + Complications + Metabolic + Age + Sex,    
                              data = .)))                

# For each of the 10 models, create tidy output
GP_nested <- GP_nested %>%
  mutate(tidy = map(.x=model, 
                    .f=broom::tidy))

# Collect the main statistics for each model 
GP_summary <- GP_nested %>%
    unnest(tidy, 
           .drop = T)

# P-values
GP_summary_pvals <- GP_summary %>% 
  filter(term!="(Intercept)") %>%
  select(GP, term, p.value) %>%
  spread(term, p.value)

# beta coefficient values
GP_summary_betavals <- GP_summary %>%
  filter(term!="(Intercept)") %>%
  select(GP, term, estimate) %>%
  spread(term, estimate)
```

# Let's make some tables

```{r}
# stargazer text table
stargazer(data.frame(GP_summary_pvals), # Treat table as a data frame for proper rounding using digits=3
          type="text",                  # Choose a text-style table (so no HTML, LaTeX...)
          summary=FALSE,                # Don't summarize the data, just print the real values
          digits=3,                     # round the numbers in the table to 3 decimals
          title="Fake data glycan peaks p-values", 
          notes="alpha = 0.05")   
```

```{r, eval=FALSE}
# stargazer html table
stargazer(data.frame(GP_summary_pvals), # Treat table as a data frame for proper rounding using digits=3
          type="html",                  # Choose a html-style table (so no HTML, LaTeX...)
          summary=FALSE,                # Don't summarize the data, just print the real values
          digits=3,                     # round the numbers in the table to 3 decimals
          title="Fake data glycan peaks p-values", 
          notes="alpha = 0.05") 
```

<table style="text-align:center"><caption><strong>Fake data glycan peaks p-values</strong></caption>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"></td><td>GP</td><td>Age</td><td>Complications1</td><td>CTR_DIAB1</td><td>Metabolic1</td><td>Sex1</td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">1</td><td>GP1</td><td>0.685</td><td>0.796</td><td>0.654</td><td>0.402</td><td>0.616</td></tr>
<tr><td style="text-align:left">2</td><td>GP10</td><td>0.593</td><td>0.055</td><td>0.495</td><td>0.891</td><td>0.062</td></tr>
<tr><td style="text-align:left">3</td><td>GP2</td><td>0.198</td><td>0.667</td><td>0.295</td><td>0.670</td><td>0.626</td></tr>
<tr><td style="text-align:left">4</td><td>GP3</td><td>0.908</td><td>0.297</td><td>0.554</td><td>0.469</td><td>0.095</td></tr>
<tr><td style="text-align:left">5</td><td>GP4</td><td>0.989</td><td>0.732</td><td>0.244</td><td>0.798</td><td>0.197</td></tr>
<tr><td style="text-align:left">6</td><td>GP5</td><td>0.552</td><td>0.008</td><td>0.567</td><td>0.067</td><td>0.909</td></tr>
<tr><td style="text-align:left">7</td><td>GP6</td><td>0.989</td><td>0.087</td><td>0.501</td><td>0.307</td><td>0.937</td></tr>
<tr><td style="text-align:left">8</td><td>GP7</td><td>0.600</td><td>0.553</td><td>0.892</td><td>0.566</td><td>0.334</td></tr>
<tr><td style="text-align:left">9</td><td>GP8</td><td>0.551</td><td>0.637</td><td>0.433</td><td>0.097</td><td>0.145</td></tr>
<tr><td style="text-align:left">10</td><td>GP9</td><td>0.331</td><td>0.824</td><td>0.425</td><td>0.140</td><td>0.866</td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td colspan="7" style="text-align:left">alpha = 0.05</td></tr>
</table>