---
title: "R Notebook"
output: html_notebook
---

## Descriptive statistics: in text format

`stargazer` will automatically recognize the type of object, and will produce the appropriate output. In the case of data frames, it will display summary statistics. For more details/options, type `?stargazer`

With `out`, the table will be saved in the working directory with whatever name you write in this option. You can open this file with any word processor.

```{r}
mydata <- mtcars
# install.packages("stargazer") # Use this to install it, do this only once
library(stargazer)
stargazer(mydata, 
          type="text", 
          title="Descriptive statistics", 
          digits=1,
          out="table1.txt")
```

```{r}
# Same output, transposed (variables in columns)
stargazer(mydata,
          type="text",
          title="Descriptive statistics",
          digits=1,
          out="table1.txt",
          flip=TRUE) # Use this option to transpose output)
```

## Descriptive statistics: in text format, replacing variable names with labels

## Regression models: various outputs

```{r}
mydata$fast <- as.numeric((mydata$mpg > 20.1))
m1 <- lm(mpg ~ hp, 
         data=mydata)
m2 <- lm(mpg ~ hp + drat, 
         data=mydata)
m3 <- lm(mpg ~ hp + drat + factor(gear), 
         data=mydata)
m4 <- glm(fast ~ hp + drat + am, 
          family=binomial(link="logit"), 
          data=mydata)

# Display table as text
stargazer(m1,
          m2,
          m3,
          m4,
          type="text")

# Display table HTML code and save as .htm file (you can open this in word or browser)
# Also change the labels of the predictors (covariate.labels option) 
# Also add a note to explain the outcome variables (dep.var.caption option)
stargazer(m1,
          m2,
          m3,
          m4,
          type="htm",
          dep.var.caption=c("Miles/(US) gallon",
                          "Fast car (=1)"),
          covariate.labels=c("Gross horsepower",
                             "Rear axle ratio",
                             "Four forward gears",
                             "Five forward gears",
                             "Type of transmission (manual=1)"),
          out="models.htm")

```


 
 
